---
title: "explore_genedf"
author: "AustinTHilliard"
date: '`r format(Sys.time(), "%d %B, %Y")`'
output: workflowr::wflow_html
---

## Setup
### Clean workspace, set options, load data
```{r start clean and load data}
rm(list=ls())
options(stringsAsFactors=FALSE)
genedf = readRDS("data/genedf.rds")

# look at the first few genes
genedf[1:5,]
```

## Brief aside about data object types and how to index them
### Note that data frames (like `genedf`) are different from matrices in R, even though they both just look like 2-d arrays
#### Data frames are usually nicer to work with but some functions require a matrix
``` {r data frames and matrices 1}
# You can check what class an object is with the class() function
class(genedf)

# You can change the class with as.matrix(), as.data.frame(), etc.
genemat = as.matrix(genedf)
class(genemat)
class(as.data.frame(genemat))
```
R objects have both a class and storage mode. The difference between class and mode can be quite confusing. If you're interested [look here](https://stackoverflow.com/questions/35445112/what-is-the-difference-between-mode-and-class-in-r), but ...



### The main differences between data frames and matrices are:
#### 1) Matrices can only contain one type of data, e.g. numbers or characters/strings

``` {r data frames and matrices 2}
# For example, genedf is a data frame that contains different types of data

```

#### 2) Matrices can only be indexed with numeric coordinates 


## Subsetting to specific genes
### One gene of interest
```{r x}
# set a specific gene name, this should be the official A. burtoni gene symbol
xgene = "ar"

# is it present at all?
any(genedf$geneSym == xgene)

# Below are examples, we'll only look at first five rows
# 1. Column number
#    Note that this is the only way to access columns in a matrix
#    
genedf[1:5, 2]

# ---

# get its row number

xgeneRow = which(genedf$geneSym == xgene)
xgeneRow
xgeneRow = which(genedf[, 2] == xgene)
xgeneRow

# these are equivalent
subset(genedf, geneSym == xgene)
genedf[xgeneRow, ]
genedf[xgeneRow, ]



```