---
title: "explore_genedf"
author: "AustinTHilliard"
date: '`r format(Sys.time(), "%d %B, %Y")`'
output: workflowr::wflow_html
---

## Setup
### Clean workspace, set options, load data
```{r start clean and load data}
rm(list=ls())
options(stringsAsFactors=FALSE)
genedf = readRDS("data/genedf.rds")

# look at the first few genes
genedf[1:5,]
```

## Subsetting to specific genes
### One gene of interest
#### Set a specific gene name
This should be the official _A. burtoni_ gene symbol. Here we will use the androgen receptor gene _ar_.
```{r}
xgene = "ar"
```
#### Is the gene present at all in `genedf`, if so, where?
`genedf$geneSym` accesses the column in `genedf` that has the name `geneSym`.
There are other ways to access columns in a data frame, but I prefer `$` notation because it's explicit.
```{r}
genedf$geneSym[1:5]
```
`genedf$geneSym == xgene` creates a vector of logical/boolean values testing whether values in `genedf$geneSym` are equal to the value in `xgene`. Entry _i_ corresponds to row _i_ in `genedf`.
```{r}
xgeneLogicalVec = genedf$geneSym == xgene
xgeneLogicalVec[1:5]
```
When a vector containing logical/boolean values is large we may want to test:
1. Whether any entries are `TRUE`
2. The exact locations of any entries that are `TRUE`
```{r}
# Note that the length of xgeneLogicalVec is the same as the number of rows in genedf
length(xgeneLogicalVec)
# Are any entries in xgeneLogicalVec TRUE?
xgenePres = any(xgeneLogicalVec)
xgenePres
# Which entries are TRUE?
xgeneRow = which(xgeneLogicalVec)
xgeneRow
```
Here are multiple equivalent ways to access the row in `genedf` containing information for the gene whose name is stored in `xgene`.
```{r}
# Use the logical vector 
genedf[xgeneLogicalVec, ]
genedf[genedf$geneSym == xgene, ]
# Use the row number
genedf[xgeneRow, ]
# Use the subset() function, which effectively creates a logical vector equivalent to xgeneLogicalVec on the fly, then uses is to subset the data frame
subset(genedf, geneSym == xgene)
```


```{r}

```



```{r}

```



```{r}

```



```{r}

```


```{r}

```


```{r}

```


```{r}

```








