---
title: "Find a gene's n nearest neighbors from a topological overlap matrix"
author: "AustinTHilliard"
date: '`r format(Sys.time(), "%d %B, %Y")`'
output: workflowr::wflow_html
---

## Setup
### Clean workspace, set options, load functions
```{r start clean}
rm(list=ls())
options(stringsAsFactors=FALSE)
source("code/NN_from_TOM_functions.R")
```

### Load and prepare data
```{r load and prepare data}
# load topological overlap matrix and gene information data frame
TOM = readRDS("data/TOM.net_small.rds")
```
_Note:_ All plots and results here were generated with a random subsampling of the full TOM, so yours will look different. 
[Download full TOM.net.rds here](https://stanford.box.com/v/networks-data-TOM-net-rds)
```{r}
genedf = readRDS("data/genedf.rds")
```
[Download genedf.rds here](https://stanford.box.com/v/networks-data-genedf-rds)
```{r}
# make sure genes are in same order in both
genedf = genedf[match(rownames(TOM), genedf$geneSym), ]

# sort genes by module and kME
new_order = order(genedf$module, -genedf$kME)

# reorder TOM and genedf
genedf = genedf[new_order, ]
rownames(genedf) = NULL
TOM = TOM[new_order,new_order]

# look at the first few genes
TOM[1:5,1:5]
genedf[1:5,]
```

## Find nearest neighbors of a gene
### Use top gene in module with highest average overexpression in ASC vs D
#### Get average dZ.ASC_D score for each module 
```{r find module with highest average dZ.ASC_D}
mod_avg_dZ.ASC_D = sort(tapply(genedf$dZ.ASC_D, genedf$module, mean))
```

#### Positive dZ.ASC_D scores reflect higher expression in ASC 
```{r, mod_avg_dZ.ASC_D barplot, fig.height=4, fig.width=9}
par(mai=c(1.5,.8,0,0))
barplot(mod_avg_dZ.ASC_D,
        col=names(mod_avg_dZ.ASC_D),
        ylab="Mean dZ.ASC_D",
        las=2)
```

#### Get gene with highest kME in royalblue module and its 25 nearest neighbors
```{r get gene with highest kME in module}
gene1 = subset(genedf, module=="royalblue")[1,]
gene1

gene1nn = .getNNeighbors(gene=gene1$geneSym, tom=TOM, n=25)
gene1nn

# make data frame with just these genes and add column with TO
gene1nn_df = as.data.frame(cbind(genedf[match(names(gene1nn), genedf$geneSym), ],
                                 TO=gene1nn))
```
<!--
```{r TO vs dZ scatterplots, warning=FALSE, fig.height=3.33, fig.width=9}
par(mfrow=c(1,3))
for (dZ in c("dZ.ASC_D","dZ.ASC_ND","dZ.D_ND")) {
  #
  y1 = gene1nn_df[, dZ]
  xcor = cor.test(gene1nn, y1, method="s")
  
  #
  plot(gene1nn, y1, 
       xlab=paste0("TO with ",gene1$geneSym), 
       ylab=dZ, 
       col=gene1nn_df$module, 
       pch=20, cex=2,
       main=paste0("rho=",signif(xcor$estimate,2),", p=",signif(xcor$p.value,2)))
  
  #
  abline(lm(y1 ~ gene1nn), lty="dashed", col="red")
}
```
-->

#### Make a circle plot of the connectivity among the genes
```{r circle plot, fig.height=6, fig.width=6}
# 
gene1nn_genes = c(as.character(gene1$geneSym), names(gene1nn))

#
xgenes = match(gene1nn_genes, rownames(TOM))
xTOM = TOM[xgenes, xgenes]

#
xlabels = gene1nn_genes
xlabels[1] = paste0(xlabels[1],"*")

#
xcolors = genedf$module[match(gene1nn_genes, genedf$geneSym)]

#
par(mai=c(1,1,1,1))
.circlePlot(adjacency=xTOM, 
           labels=xlabels, 
           order=order(-apply(xTOM,1,sum)),
           lineColors=.grey2red(100, 1, 1),
           pointBg=xcolors,
           min.cex.points=1, max.cex.points=2,
           min.cex.labels=.6, max.cex.labels=1.2)
```
<!--
### Both copies of androgen receptor
#### Find ids for both genes and get their 20 nearest neighbors
```{r find ARs}
ARdf = subset(genedf, grepl('androgen receptor', desc, ignore.case=T))
ARdf

arNN = lapply(as.list(ARdf$geneSym), 
              function(g) .getNNeighbors(gene=g, tom=TOM, n=20))
names(arNN) = ARdf$geneSym
arNN
```

```{r}
arNN_dfs = lapply(arNN, 
                  function(g) genedf[match(names(g), genedf$geneSym), ])
```

#### Make circle plots
```{r AR circle plots, fig.height=12, fig.width=6}
# gene1nn_genes = c(as.character(gene1$geneSym), names(gene1nn))
# 
arNNgenes = lapply(arNN, names)
par(mfrow=c(2,1), mai=c(1,1,0,1))
for (g in 1:2) {
  arNNgenes[[g]] = c(names(arNNgenes)[g], arNNgenes[[g]])
  xgenes = match(arNNgenes[[g]], rownames(TOM))
  xTOM = TOM[xgenes, xgenes]
  xlabels = arNNgenes[[g]]
  xlabels[1] = paste0(xlabels[1],"*")
  xcolors = genedf$module[match(arNNgenes[[g]], genedf$geneSym)]

  .circlePlot(adjacency=xTOM,
             labels=xlabels,
             order=order(-apply(xTOM,1,sum)),
             lineColors=.grey2red(100, 1, 1),
             pointBg=xcolors,
             min.cex.points=1, max.cex.points=2,
             min.cex.labels=.6, max.cex.labels=1.2)
}
```

-->